Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-09-04 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:09:03	15:17:02	(Thu 3.Sep)
la2pdf commit - 2015:09:04	12:58:22	(Fri 4.Sep)

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/mvdr_selected_windows_holmengraa.pdf}%
\caption{\emph{Determining LCA window type}: MVDR images with typical frequency responses for windows used in various pixel regions.\newline
\emph{Left:}\hfill
\parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, using data collected from the 32 element HISAS1030 sonar attached to the HUGIN AUV. MVDR was run with $K=1$ temporal oversampling, $L=16$ subarray size, $d=1\%$ diagonal loading and 8 times lateral oversampling.}\protect\\\hspace{\textwidth}
\emph{Right:}\hfill
\parbox[t]{.95\linewidth}{MVDR window responses computed from 40x40 pixel groups in the shadow, highlight and speckle region of the Holmengraa image. The responses are overlayed each other. The dashed red lines at -13dB and -43dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3dB in highlight. The phase varies mostly in highlight to allow the steering there.} }\label{mvdr_selected_windows}
\end{figure*}

@@

\begin{figure*}[t]%\centering%
\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds.pdf}%
\caption{\emph{Determining Kaiser parameter boundaries}: LCA images created with a large and equal size window database, but where the upper bounds for $\beta$ and $\phi$ is varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the distance from the maximum to the -3dB point of the anchor
\newline
\emph{Upper bound $\beta$:}\hfill
\parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{mvdr_selected_windows} suggests that MVDR tend to choose windows with levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
\emph{Upper bound $\phi$:}\hfill
\parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
}\label{oversampling_mosaic_bounds}
\end{figure*}

@@

Adding slightly steered versions of each window to the window database gives LCA greater flexibility in searching for optimal window. A Kaiser window $\f_\beta$ steered to $\phi$ can be expressed as 
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}
\end{align}
%
where diag($\a_\phi$) is a diagonal matrix constructed from the steering vector $\a_\phi$:
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{md}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{(M-1)d}{\lambda}\sin(\phi)}
}.
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_definition}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

We call the upper bound for steering $\phi_{\text{-3dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a -3dB crossing that is exactly half that of the unsteered window. Using a numerical search for the steering angle $\phi$ that fulfilled this criterion for a wide range of $\beta$-values, we obtained a third degree polynomial fit to these points:
%
\begin{align}
\phi_\text{3dB}(\beta) \approx \minus{}0.002894\beta^3 + 0.02451\beta^2 - 0.01308\beta + 0.2980
\end{align}
%
Henceforth any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

@@

Need $M$ in the image sector. Why???

- Abs -> adds a factor 2
- log -> infinity?
- near field -> range and azimuth is coupled

@@

While MVDR is free to choose non-symmetric windows, we can observe from \Fig{mvdr_selected_windows} that it prefers symmetric ones. In shadow and speckle the windows are fully symmetric, while in highlight the windows are approximately symmetric within the illuminated\todo{valid term?} sector of the seafloor. The peak sidelobe levels are between that of the rectangular and Hamming window. This corresponds to Kaiser windows with $\beta$-values of 0 and 2, respectively.

@@

\Fig{oversampling_mosaic_bounds} show LCA images made with the same sized window database, but containing windows with varied upper boundaries for $\beta$ and $\phi$. We suggest an upper bound for $\beta$ between 2 and 5, where 5 is slightly better at suppressing sidelobes. For $\phi$ we suggest to steer either to 50\% or 100\% of the -3dB width, which leads to a resolution of 73\% or 54\% of the rectangular window, respectively. As lower boundaries we choose $\beta=0$ to always include a rectangular window, and $\phi=0$ to always include the centered window.

>>>>

\begin{figure*}[t]\centering%
\includegraphics[width=\linewidth]{gfx/mvdr_selected_windows_holmengraa.pdf}%
\caption{\emph{Determining LCA window type}: MVDR image with typical frequency responses for windows used in various pixel regions.\newline
\emph{Left:}\hfill
\parbox[t]{.95\linewidth}{MVDR sectorscan image of the oiltanker Holmengraa, using data collected from the 32 element HISAS1030 sonar attached to the HUGIN AUV. MVDR was run with $K=1$ temporal oversampling, $L=16$ subarray size, $d=1\%$ diagonal loading and 8 times lateral oversampling.}\protect\\\hspace{\textwidth}
\emph{Right:}\hfill
\parbox[t]{.95\linewidth}{MVDR window responses computed from 40x40 pixel groups in the shadow, highlight and speckle region of the Holmengraa image. The responses are overlayed each other. The dashed red lines at -13dB and -43dB marks the peak sidelobe levels of an unsteered rectangular and Hamming window, respectively. Note how the responses are more or less symmetric, with very little steering in the shadow, moderate steering in speckle and steering within roughly 3dB in highlight. The phase varies mostly in highlight to allow the steering there.} }\label{mvdr_selected_windows}
\end{figure*}

@@

\begin{figure*}[t]%\centering%
\includegraphics[width=\textwidth]{gfx/oversampling_mosaic_bounds.pdf}%
\caption{\emph{Determining Kaiser parameter boundaries}: LCA images created with a large and equal size window database, but where the upper bounds for $\beta$ and $\phi$ is varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two -3dB points of the anchor. These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
\newline
\emph{Upper bound $\beta$:}\hfill
\parbox[t]{.89\linewidth}{Observe that LCA gets better at suppressing sidelobes in the image as we increase the upper bounds for $\beta$. \Fig{mvdr_selected_windows} suggests that MVDR tend to choose windows with levels lower than that of Kaiser with $\beta=2$, but here we observe further improvement going to $\beta=5$.}\newline
\emph{Upper bound $\phi$:}\hfill
\parbox[t]{.89\linewidth}{As we increase the upper bound of the steering $\phi$, we also increase the lateral image resolution. }
}\label{oversampling_mosaic_bounds}
\end{figure*}

@@

Adding slightly steered versions of each window to the window database gives LCA greater flexibility in searching for optimal window. A Kaiser window $\f_\beta$ steered to the angle $\phi$ can be expressed as 
%
\begin{align}
\w_{\beta,\phi} = \frac{\f_\beta\H\diag{(\a_\phi)}}{\f_\beta\H\a_\phi}
\end{align}
%
where diag($\a_\phi$) is a diagonal matrix constructed from the steering vector $\a_\phi$:
%
\begin{align}
\a_\phi = \bmat{
1 \\
e^{-j\frac{md}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{(M-1)d}{\lambda}\sin(\phi)}
}.
\end{align}
%
Here $d$ is the element spacing, $\lambda$ is the wavelength and $\phi$ is the steering amount. The normalization factor $\f_\beta\H\a_\phi$ is the reciprocal of the window's coherent gain. It ensures unit gain in the direction of interest as required by (\ref{eq_lca_definition}). Since the signal-to-noise ratio is constant for a specific window (same $\beta$-value), this normalization also proportionally increases the incoherent noise gain.

@@

We call the upper bound for steering $\phi_{\text{3dB}}(\beta)$, and define it as the steering angle needed for the steered response to have a -3dB crossing that is exactly half that of the unsteered window. Using a numerical search for the steering angle $\phi$ that fulfilled this criterion for a wide range of $\beta$-values, we obtained a third degree polynomial fit to these points:
%
\begin{align}
\phi_\text{3dB}(\beta) \approx \minus{}0.002894\beta^3 + 0.02451\beta^2 - 0.01308\beta + 0.2980
\end{align}
%
Henceforth any steering will be specified relative to this upper bound. With each window steered this way we expect the resolution gain to remain predictable and independent of $\beta$, and we also effectively constrain the white noise gain of the beamformer.

@@

\todo{Need $M$ in the image sector. Why???}

\todo{
- Abs -> adds a factor 2\\
- log -> infinity?\\
- near field -> range and azimuth is coupled}

@@

While MVDR is free to choose non-symmetric windows, we can observe from \Fig{mvdr_selected_windows} that it prefers symmetric ones. In shadow and speckle the windows are fully symmetric, while in highlight the windows are approximately symmetric within the illuminated\todo{valid term?} sector of the seafloor. The peak sidelobe levels are between that of the rectangular and Hamming window.

All the responses in \Fig{mvdr_selected_windows} can be approximated with the Kaiser-Bessel function. By adjusting the window parameter $\beta$ we can adjust the trade-off between mainlobe width and sidelobe level. 
To generate windows with different mainlobe width and sidelobe level,

A parametric window like Kaiser-Bessel allow us to create approximately these and the responses in between.

The Kaiser-Bessel windows s with $\beta$-values of 0 and 2 allows us to approximately trade between these two boundaries. \todo{add the advantages here?}

In our studies

@@

Letting LCA choose from Kaiser windows seems like a good choice due to their similarity to the 
Which Kaiser windows should we include in the windows database? The parameter $\beta$ controls the window shape, and $\phi$ the steering.

Having chosen to fill the window database with steered Kaiser windows, we should  function While having decided to use steered Kaiser windows, we still  To decide what the ideal boundaries for $\beta$ and $\phi$ should be, we constructed large and equally sized window databases containing windows with different parameter boundaries. The images produced are shown in \Fig{oversampling_mosaic_bounds}. To include and compare with a rectangular weighted DAS, we set the lower boundary to 0 for both parameters.

varied upper boundaries for $\beta$ and $\phi$.

 The upper left image has $\beta=0$ and no steering, and is thus equal to a rectangularly weighted DAS.

We suggest an upper bound for $\beta$ between 2 and 5, where 5 is slightly better at suppressing sidelobes. For $\phi$ we suggest to steer either to 50\% or 100\% of the -3dB width, which leads to a resolution of 73\% or 54\% of the rectangular window, respectively. As lower boundaries we choose $\beta=0$ to always include a rectangular window, and $\phi=0$ to always include the centered window.

LCA images created with a large and equal size window database, but where the upper bounds for $\beta$ and $\phi$ is varied. The upper left image is equal to DAS with a rectangular window. For all images we measure the lateral distance $\Delta x$ between the two -3dB points of the anchor. These are specified relatively to the reference distance  $\Delta x_\text{ref}$ from the upper left DAS image.
